Technical Dimensions of Programming Systems

### 2022-05-18 prog-2022/summary-table.tex
Add factoring of complexity and rename automation to complexity

\begin{table}
\begin{tabular}{ >{\raggedleft\arraybackslash}p{0.2\linewidth}  p {0.4\linewidth}  >{\raggedright\arraybackslash}p{0.4\linewidth} }
\hline
Dimension (CLUSTER) & Summary & Range of key examples \\ \hline
\hline CUSTOMIZABILITY &
\summ{Once a program exists in the system, how can it be extended and modified?}
Staging of Customization &
Must we customize \emph{running} programs differently to \emph{inert} ones? Do these changes last beyond termination? &
Source code vs. config files, Developer Tools tab, auto image-based persistence, scripting language \\
Externalizability &
Which portions of the system's state can be referenced and transferred to/from it? &
None (state is private) vs. all state exposed as human-legible, CSS-like addressing \\
Additive Authoring &
How far can the system's behavior be changed by \emph{adding} expressions? &
None (requires power to change original) vs. full (anything can be overridden repeatedly) \\
Self-Sustainability &
How far can the system’s behavior be changed from within? &
None (rely on extenal tools) vs. self-sufficient (contains everything needed) \\
\hline AUTOMATION &
\summ{How far does the system remove the need to spell out implementation in minute detail?}
Degrees of Automation &
What part of program logic does not need to be explicitly specified? &
Garbage collection (low-tech) vs. Prolog engine (hi-tech) \\
\hline ERRORS &
\summ{What does the system consider to be an error; how are they prevented and handled?}
Error Detection &
What errors can be detected in which feedback loops, and how? &
Human inspection in live coding vs. partial automation in static typing \\
Error Response &
How does the system respond when an error is detected? &
Does it stop, recover automatically, ignore the error or ask the user how to continue? \\
\hline ADOPTABILITY &
\summ{How does the system facilitate or obstruct adoption by both individuals and communities?}
Learnability &
What is the attitude towards the \emph{learning curve} and what is the target audience? &
HyperCard for the general public vs. FORTRAN for scientists \\
Sociability &
What are the social and economic factors that make the system the way it is? &
Funding, volunteers, code sharing, Q/A sites vs. documentation, sense of belonging \\
\hline
\end{tabular}
\caption{\label{summary-table2} Quick reference sheet for our set of Technical Dimensions.}
\end{table}

>>>>

\begin{table}
\begin{tabular}{ >{\raggedleft\arraybackslash}p{0.2\linewidth}  p {0.4\linewidth}  >{\raggedright\arraybackslash}p{0.4\linewidth} }
\hline
Dimension (CLUSTER) & Summary & Range of key examples \\ \hline
\hline CUSTOMIZABILITY &
\summ{Once a program exists in the system, how can it be extended and modified?}
Staging of Customization &
Must we customize \emph{running} programs differently to \emph{inert} ones? Do these changes last beyond termination? &
Source code vs. config files, Developer Tools tab, auto image-based persistence, scripting language \\
Externalizability &
Which portions of the system's state can be referenced and transferred to/from it? &
None (state is private) vs. all state exposed as human-legible, CSS-like addressing \\
Additive Authoring &
How far can the system's behavior be changed by \emph{adding} expressions? &
None (requires power to change original) vs. full (anything can be overridden repeatedly) \\
Self-Sustainability &
How far can the system’s behavior be changed from within? &
None (rely on extenal tools) vs. self-sufficient (contains everything needed) \\
\hline COMPLEXITY &
\summ{How does the system structure complexity and what level of detail is required?}
Factoring of Complexity &
What programming details are hidden in reusable components and how? &
Domain-specific (more hiding) vs. general-purpose (less hiding) programming \\
Level of Automation &
What part of program logic does not need to be explicitly specified? &
Garbage collection (low-tech) vs. Prolog engine (hi-tech) \\
\hline ERRORS &
\summ{What does the system consider to be an error; how are they prevented and handled?}
Error Detection &
What errors can be detected in which feedback loops, and how? &
Human inspection in live coding vs. partial automation in static typing \\
Error Response &
How does the system respond when an error is detected? &
Does it stop, recover automatically, ignore the error or ask the user how to continue? \\
\hline ADOPTABILITY &
\summ{How does the system facilitate or obstruct adoption by both individuals and communities?}
Learnability &
What is the attitude towards the \emph{learning curve} and what is the target audience? &
HyperCard for the general public vs. FORTRAN for scientists \\
Sociability &
What are the social and economic factors that make the system the way it is? &
Funding, volunteers, code sharing, Q/A sites vs. documentation, sense of belonging \\
\hline
\end{tabular}
\caption{\label{summary-table2} Quick reference sheet for our set of Technical Dimensions.}
\end{table}
